\documentclass[journal]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{url}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{bm}
\usepackage{gensymb}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, fit, backgrounds}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Math shortcuts
\newcommand{\vR}{\mathbf{R}}
\newcommand{\vp}{\mathbf{p}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\va}{\mathbf{a}}
\newcommand{\vom}{\boldsymbol{\omega}}
\newcommand{\vg}{\mathbf{g}}
\newcommand{\vu}{\hat{\mathbf{u}}}
\newcommand{\vb}{\mathbf{b}}
\newcommand{\vz}{\mathbf{z}}
\newcommand{\vOm}{\boldsymbol{\Omega}}
\newcommand{\SO}{\mathrm{SO}(3)}
\newcommand{\so}{\mathfrak{so}(3)}
\newcommand{\Exp}{\mathrm{Exp}}
\newcommand{\Log}{\mathrm{Log}}
\newcommand{\calX}{\mathcal{X}}

% --- Cross-references: \autoref with IEEE-style names ---
\usepackage[hidelinks]{hyperref}
\renewcommand{\sectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Section}
\renewcommand{\subsubsectionautorefname}{Section}
\renewcommand{\figureautorefname}{Fig.}
\renewcommand{\tableautorefname}{Table}
\renewcommand{\equationautorefname}{Eq.}

% --- Acronyms: first \ac{} use writes out the long form + short; later
%     uses print the short form only (\printacronyms lists them). ---
\usepackage{acro}
\acsetup{list/template=tabular}  % auto-size the short-form column
\input{acronyms}

% Modestly tighten float/text separation to keep the bibliography within
% the 12-page limit (reversible; no content change).
\setlength{\textfloatsep}{8pt plus 2pt minus 2pt}
\setlength{\floatsep}{6pt plus 2pt minus 2pt}
\setlength{\intextsep}{6pt plus 2pt minus 2pt}

\begin{document}

\title{Continuous-Time Radar-Inertial Odometry\\with B-Spline Sliding-Window Optimization}

\author{Timo~Weiss~and~Lukas~Pries%
\thanks{Manuscript received Month~XX,~2026; revised Month~XX,~2026.}%
\thanks{The authors are with the Department of Informatics, Technical
University of Munich (TUM), 80333 Munich, Germany (e-mail:
timo.weiss@tum.de; lukas.pries@tum.de).}%
\thanks{Code and datasets are available at
\protect\url{https://github.com/Dr1zz3l/radar-iwr6843-driver}.}}

\markboth{IEEE Transactions on Robotics,~Vol.~XX, No.~X, Month~2026}%
{Weiss and Pries: Continuous-Time Radar-Inertial Odometry}

\maketitle

%------------------------------------------------------------------
\begin{abstract}
We present a radar-inertial odometry (RIO) system for 6-DOF state
estimation on a quadrotor using a single-chip 60\,GHz mmWave radar
(10--60 points per frame) and a 1\,kHz IMU, evaluated through aggressive
racing and attitude-tumbling flight up to \SI{10}{\radian\per\second}
--- a regime existing RIO evaluations do not cover, where the radar
model itself degrades.  The trajectory is a continuous-time quintic
position B-spline and a cumulative SO(3) orientation B-spline, optimized
by a fixed-lag smoother with Schur complement marginalization at a
0.3\,s stride over a 3\,s window; a sensor-only pipeline initializes it
without external \emph{pose} reference.  Doppler measures velocity
directly, so we evaluate RIO as a \emph{velocity and orientation}
source, reporting position as drift per distance.  A Markov-blanket
factor-selection rule makes the marginalization a true marginal, and a
single dynamics-adaptive measurement-weighting law covers hover to
\SI{10}{\radian\per\second}, validated on held-out flights.  At the live
leading edge, velocity and orientation RMSE stay below
\SI{0.5}{\metre\per\second} and \SI{3}{\degree} across racing (position
drift under \SI{1}{\percent}) and degrade gracefully to
\SI{6.4}{\degree} orientation (\SI{2.9}{\percent} drift) under tumbling;
the output
orientation covariance is approximately calibrated (NEES) on the racing
flights.  We also document negative
results: a diagnostic for marginalization inconsistency, and that the
tumbling elevation bias is a flight-regime proxy, not an antenna
constant.
\end{abstract}

\begin{IEEEkeywords}
radar-inertial odometry, continuous-time estimation, B-spline,
sliding window, marginalization, mmWave radar, Doppler velocity
\end{IEEEkeywords}


% ---- Body: one file per section (sections/) ----
\input{sections/introduction}
\input{sections/related_work}
\input{sections/system_overview}
\input{sections/methodology}
\input{sections/experimental_setup}
\input{sections/results}
\input{sections/limitations}
\input{sections/conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

% Author biographies (<=100 words each; required for the final accepted version,
% optional for initial submission). Omitted in the initial submission to respect
% the 12-page limit; uncomment for the accepted version (add degrees/dates/photos).
% \begin{IEEEbiographynophoto}{Timo Weiss}
% is with the Technical University of Munich (TUM), Munich, Germany. His
% research interests include radar-inertial odometry, continuous-time
% state estimation on Lie groups, and aerial robotics in
% perception-degraded conditions.
% \end{IEEEbiographynophoto}
%
% \begin{IEEEbiographynophoto}{Lukas Pries}
% is with the Technical University of Munich (TUM), Munich, Germany, where
% he supervises research in radar perception and autonomous aerial systems.
% \end{IEEEbiographynophoto}

\end{document}

