\section{Conclusion}
\label{sec:conclusion}

\emph{Can a single-chip mmWave radar plus \ac{imu} provide usable 6-DoF
state estimation?}  Yes --- as a \emph{velocity and orientation}
source: at the live leading edge,
\SI{0.48}{\metre\per\second}/\SI{2.0}{\degree} on gentle racing,
\SI{0.32}{\metre\per\second}/\SI{2.9}{\degree} on aggressive racing
(0.35--0.70\,s per window on a laptop), confirmed on held-out flights
and with an approximately calibrated (\ac{nees}) orientation covariance
on racing.
Absolute position is not observable from Doppler+\ac{imu} and drifts with
distance (0.6--0.9\% on racing, 2.9\% under tumbling); yaw requires a
heading sensor.  Within those structural limits, \emph{good results
took three things}: (1) a \emph{consistent} fixed-lag smoother ---
the Markov-blanket factor-selection rule makes the marginalization a
true marginal, eliminating regime-dependent prior tuning whose
symptom pattern we document as a reusable diagnostic;
(2) a \emph{dynamics-adaptive trust allocation} across the
measurement streams (rate-adaptive gyroscope weight, rate-inflated
radar/accelerometer noise, per-point \ac{snr} weighting) --- one
configuration from hover to \SI{10}{\radian\per\second}, with the
velocity--orientation trade made explicit as a measured Pareto curve;
and (3) \emph{characterized sensor systematics} --- including the
finding that the apparent elevation bias under tumbling is a
flight-regime error proxy, not an antenna constant, and that the
extrinsic pitch is weakly observable, best fixed at its measured value
rather than estimated online.  Notably, the largest accuracy gains came from diagnosis-led
re-weighting of existing measurements, not from added model capacity:
a representation-level study showed the orientation spline already
tracks \SI{10}{\radian\per\second} flips at the gyroscope noise floor,
disproving our own adaptive-knot-density hypothesis before
implementation.

Open limitations: (1) loop-scale position geometry through flips is
bounded by during-flip radar data quality (Doppler noise
$15\times$ the level-flight core), demonstrated by an asymmetric
factor-split ablation; (2) ground truth itself degrades during flips
(\ac{mocap} occlusion), bounding what can be measured there; and
(3) cross-validation against the \ac{ekf}-\ac{rio} family
(\autoref{sec:baselines}) reveals a design-space split.  With our
default \ac{ransac} front-end --- the single-chip outlier rejection
\mbox{reve} uses --- our solver reaches the same order of magnitude on
position on their public flights ($0.5$--$1.9$\,m, trailing the
yaw-aided baseline), matches their roll/pitch, and leads on heading-aided
orientation; the prefilter closes the elevation-bias gap that otherwise
leaves a Huber-only port at $12$--$15\%$ vertical drift.  In reverse,
their stock filters, starved by our sparse ${\sim}13$-point clouds and
racing-frame vibration, reject up to 99.7\% of radar updates and
dead-reckon on slow racing.  Each stack is engineered for its own regime
--- dense clouds and a low-vibration carrier versus sparse, finely
quantized Doppler on an aggressive platform.
