\section{Limitations and Negative Results}
\label{sec:limits}

\subsection{Statistical Scope}
\label{sec:stat_scope}
Each flight regime is evaluated on a \emph{single} representative
recording, so the headline numbers carry no run-to-run error bars.  The
held-out flights (\autoref{sec:heldout}), the cross-platform validation
(\autoref{sec:baselines}), and the deterministic, bit-identical re-runs
broaden the evidence, but they do not substitute for repeated trials:
re-runs add reproducibility, not statistical power.  Reporting
mean\,$\pm$\,std over several recordings per regime is the natural
strengthening, and is left to future data collection.

\subsection{Heading Sensor Requirement}
\label{sec:yaw}
Yaw is a gauge freedom for Doppler: rotating the entire trajectory
about the world gravity axis leaves all radial velocity measurements
unchanged, so absolute heading cannot be recovered from radar and \ac{imu}
alone.  A magnetometer resolves this and is standard on virtually all
flight platforms; here \ac{mocap} yaw serves that role.  Because that
\ac{mocap} yaw is noise- and bias-free (unlike a real magnetometer),
the heading-aided orientation numbers are in principle optimistic.  We
bound this by re-running with a magnetometer surrogate on the heading
prior: a slowly-varying $2\degree$ heading drift (a half-period sine over
the flight, so it is not a gauge the start-anchored alignment cancels)
plus $1.5\degree$ white noise, seeded.  Live orientation degrades by at
most $0.3\degree$ (fast racing $2.88\to3.17\degree$) and is essentially
unchanged on slow racing ($1.97\to1.88\degree$) and backflips
($6.35\to6.34\degree$): the \SI{1}{\kilo\hertz} gyroscope constrains yaw
\emph{rate} so the slow drift is resisted, the white noise is averaged by
the prior, and backflip orientation error is dominated by flip-time
roll/pitch, not yaw.  The headline orientation is therefore robust to
heading-sensor imperfection, and the heading-\emph{independent} roll/pitch
decomposition (\autoref{tab:baselines}(b)) and the unaided-yaw check
(\autoref{sec:baselines}) further isolate attitude from this idealization.
This is a sensor-architecture requirement, not an open research problem.

\subsection{Extreme Dynamics: Backflips}
\label{sec:backflips}

\textbf{Batch.}
The batch solver handles backflips only with $\Delta t_o {=} 0.8$\,ms
(the \emph{batch} density, 10$\times$ denser than the \SI{8}{\milli\second}
sliding-window grid of \autoref{tab:config}; 22{,}630 knots over 18\,s) and
locked extrinsics (free pitch drifts to $+18\degree$ due to dense
under-constrained orientation DoF).  At this density the batch
orientation is \emph{bistable} (detailed below): under small
timestamp perturbations it oscillates between ${\sim}8\degree$ and
${\sim}25\degree$ regimes, so we do not report a single batch figure and
take the sliding-window result as the trustworthy one.

\textbf{Sliding window.}
The fixed-lag smoother initially appeared fundamentally unsuited to
backflips: with the inconsistent prior and a (silently inactive)
position anchor, the best result was 2.53\,m/10.8\degree\ settled.
A chain of three fixes revised this verdict.
(i)~The consistent unscaled prior (\autoref{sec:sw}) restores
inter-window information transfer.
(ii)~A soft position anchor to the dead-reckoned initialization
($\lambda_{\mathrm{pos\text{-}init}}{=}0.5$ per CP) bounds the
position random walk that sparse, rate-inflated radar cannot prevent
within a 3\,s window (without it position diverges to 8\,m; on racing
bags the same anchor is \emph{harmful}: a flip-regime rescue,
not a general ingredient).
(iii)~The dynamics-adaptive weighting law (\autoref{sec:models}),
in particular rate-adaptive gyroscope trust, which on this bag alone
is worth $\sim$3\degree\ of orientation at zero position cost,
plus the flip-regime Doppler correction $b{=}{-}1.5$\,m/s.
Together these reach \textbf{1.67\,m\,/\,5.0\degree\ settled and
1.55\,m\,/\,6.4\degree\ live at $\sim$0.6\,s per window}
(\autoref{tab:sw}; \autoref{fig:traj}).  The default \ac{ransac}
prefilter (\autoref{sec:models})
is neutral here (backflip frames are already sparse and mostly fall
below its five-return floor, so they bypass it), and the small change
from the Huber front-end ($1.51\to1.55$\,m live) is within run-to-run
noise.

\emph{Why $\Delta t_o{=}8$\,ms in the window.}
At the batch density $\Delta t_o{=}0.8$\,ms a 3\,s window has
$\sim$11{,}250 orientation DoF against $\sim$9{,}000 gyroscope
constraints (1.26 DoF/constraint); the per-window Jacobian is
rank-deficient and the solver freezes near its initialization.
At 8\,ms the window is well-conditioned (62.5\,Hz model bandwidth
comfortably covers the 10--20\,Hz flip dynamics).
The same density is what makes the \emph{batch} problem bistable (the
$\sim$8\degree/$\sim$25\degree\ oscillation noted above): the final cost
\emph{anti-correlates} with accuracy (the underconstrained problem
overfits the wrong global orientation), so single-run comparisons at
this density are not trustworthy.

\emph{Attributing the z-axis failure, and what $b$ really is.}
A visual trajectory check (numeric \ac{rmse} conceals it) showed the SW
estimate sinking in altitude.  Disabling the position anchor exposes
the raw behaviour: a near-constant $-0.57$\,m/s world-$z$ drift
($-8$\,m over the flight).  Sweeping the per-point correction
$v_{\mathrm{corr}} = v - b\,u_{s,z}$ produces a linear family of
$z$-error curves and large negative $b$ flattens the drift, so we
initially interpreted $b$ as the antenna-fixed elevation bias of the
2-TX array.  A joint calibration experiment \emph{disproves} this
interpretation: (i)~with extrinsic pitch \emph{locked} at its
measured value, level-flight racing degrades catastrophically under
any explicit $b$ (fast-racing live position
0.39\,$\to$\,3.2\,$\to$\,6.6\,m for $b{=}0,-0.5,-1.0$): a true
antenna-fixed bias would be tolerated, since nothing else can absorb
it; and (ii)~on backflips the optimum keeps growing monotonically
past the \ac{wls}-measured static bias of $-0.5$\,m/s (through $-1.5$
to $-2.0$), into physically implausible territory.
We conclude $b$ is a \emph{flight-regime error proxy}: it absorbs
an attitude-correlated Doppler systematic during flips (most likely
intra-frame motion distortion) that happens to project like an
elevation term, and we cap it at $b{=}{-}1.5$ rather than chase
the monotone trend without a mechanism.
Relatedly, extrinsic pitch is a \emph{plateau}: over a $3{\times}3$ grid
over (locked pitch, $b$) the \ac{rmse} is indistinguishable across
26.5--28.5\degree\ on backflips, and locking at 27.5\degree\ versus
online sliding-window self-calibration is \ac{rmse}-neutral on racing.  We
freeze it rather than estimate it online because free pitch is only
weakly observable and \emph{init-dependent in both} the windowed solve
and the full-trajectory batch: the converged value tracks the
initialization (a $25\degree$ seed gives $27\degree$, a $30\degree$ seed
gives $36\degree$ on both racing bags, diverging from a saddle near
$23\degree$).  The earlier ``self-calibrated $+2\degree$ absorbs the
elevation bias'' reading is retired; we instead fix pitch at the
physically measured as-built angle ($27.5\degree$, inclinometer;
\autoref{sec:results}), a few degrees below the $30\degree$ \ac{cad} design
(the 3D-printed, loosely-fastened fixture accounts for the difference).

\emph{Remaining gap.}
The residual $\sim$6\degree\ live orientation error during flips
attenuates the reconstructed loop circles relative to ground truth.
This is not a spline-bandwidth limit (\autoref{sec:negative}: denser
knots do not help); the residual lever is the radar measurement model,
whose intra-frame Doppler smearing during flips is best addressed by
\emph{per-chirp timestamps} rather than knot density, which we leave to
future work.


\subsection{Prior Scale Sensitivity as a Symptom of Inconsistent
Marginalization}
\label{sec:prior_scale}

Our initial marginalization implementation selected all residuals
touching marginalized \emph{or} boundary variables, which, through
the shared bias block, baked every \ac{imu} factor of the window into the
Schur complement, conditioned on interior states, and double-counted
factors re-added in the next window (\autoref{sec:sw}).  The
resulting overconfident prior had to be deflated by a hand-tuned
scale, and the required scale varied by orders of magnitude across
flight regimes with a harmful intermediate region.  We document this
behaviour because the symptom pattern (no universal prior weight,
non-monotone sensitivity, robust losses on the prior acting backwards)
is a useful \emph{diagnostic for marginalization inconsistency} in
sliding-window estimators generally.
\autoref{tab:prior_scale} shows the sweep measured with the
inconsistent prior.

\begin{table}[t]
\centering
\caption{Prior Scale Sensitivity (SW, live-edge \ac{rmse})}
\label{tab:prior_scale}
\small
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}lcccc@{}}
\toprule
\texttt{scale} & \multicolumn{2}{c}{Slow racing} & \multicolumn{2}{c}{Fast racing} \\
 & Pos (m) & Ori (\degree) & Pos (m) & Ori (\degree) \\
\midrule
$2{\times}10^{-4}$                       & 0.543 & 2.16 & \textbf{0.825} & \textbf{3.64} \\
$10^{-5}$                                & ---   & ---  & 0.942 & 3.58 \\
$10^{-6}$                                & 0.442 & 2.21$^*$ & 0.951 & 3.57 \\
$\bm{10^{-7}}$ \textbf{(slow default)}  & \textbf{0.338} & \textbf{2.08} & 1.279 & 3.57 \\
$10^{-8}$                                & 0.336 & 2.08 & 1.344 & 3.57 \\
\bottomrule
\multicolumn{5}{@{}p{\columnwidth}@{}}{\scriptsize $^*$Orientation
  local maximum: 2.21\degree\ $>$ 2.16\degree\ (2e-4) and
  2.08\degree\ (1e-7).  ---: scale not evaluated for this bag.}\\
\multicolumn{5}{@{}p{\columnwidth}@{}}{\scriptsize Measured under the
  inconsistent-prior diagnostic configuration (predates the \ac{ransac}
  default); the sensitivity \emph{pattern}, not the absolutes, is the
  result.}
\end{tabular}
\end{table}

Slow racing is best served by a near-zero prior ($10^{-7}$): with
gentle dynamics, each 3\,s window is self-sufficient and additional
inter-window coupling is harmful.  Fast racing instead needs the prior
($2{\times}10^{-4}$) for inter-window position continuity;
without it (scale$=10^{-7}$) fast-racing live position degrades from
0.825\,m to 1.279\,m.  The two bags thus want opposite scales, and the
non-monotone slow-racing orientation maximum at $10^{-6}$
(\autoref{tab:prior_scale}) shows no single scale is universally optimal.

The information asymmetry in the inconsistent $S$ (orientation
eigenvalues $\sim$$5.5{\times}10^{14}$, position $\sim$$10^4$) meant
no single scalar could address both modalities; eigenvalue clipping
and Cauchy robustification also failed (Cauchy down-weights the bag
that needs the prior most, since aggressive dynamics produce larger
prior residuals).

\textbf{Resolution.}
With the Markov-blanket factor-selection rule of \autoref{sec:sw},
the prior is a true marginal: a single \emph{unscaled} prior
(scale $=1$) is optimal or near-optimal on all three bags
simultaneously, the prior's Mahalanobis residual at each solution is
$\mathcal{O}(1)$, and the entire per-mission tuning table becomes
obsolete.  The fix also improves accuracy: fast-racing settled
position improves from 0.727\,m (tuned inconsistent prior) to
0.701\,m (unscaled consistent prior).  Both figures isolate the
prior-consistency effect at a fixed pre-\ac{ransac} operating point; the
\ac{ransac} default brings the same metric to 0.312\,m
(\autoref{tab:sw}).  Stride-boundary position
jumps in the settled trajectory also disappear (settled velocity \ac{rmse}
0.89\,$\to$\,0.21\,m/s on slow racing).

\subsection{Other Negative Results}
\label{sec:negative}

\textbf{Adaptive knot density.}
We tested whether denser or non-uniform orientation knots reduce the
residual backflip attitude error; they do not.  At $\Delta t_o{=}8$\,ms
the orientation spline already represents the MoCap attitude to
1.28\degree\ \ac{rmse} \emph{during} the flips, and its angular-velocity
tracking of the gyro stream is 0.295\,rad/s RMS, flat across
$\Delta t_o\in\{16,8,4\}$\,ms and sitting at the rate-independent
vibration floor (0.17\,rad/s).  The flip orientation gap is therefore
\emph{not} a spline-bandwidth limit: the spline can represent the
motion, but the optimizer does not rotate to it (the gyro residual
correlates with $|\vom_b|$ at $+0.94$ against the \emph{solved}
trajectory, and open-loop gyro dead-reckoning is 7.8\degree, better
than the 9.2\degree\ this fusion reaches \emph{at the diagnostic
configuration}, before the rate-adaptive gyro trust of
\autoref{sec:backflips} brings it to the \SI{6.4}{\degree} headline
of \autoref{tab:sw}).  The lever is measurement weighting and
robustness (\autoref{sec:backflips}), not knot density: the
dynamics-adaptive gyro trust closes most of the gap, \emph{disproving}
the adaptive-knot-density hypothesis for this regime and obviating the
non-uniform-basis extension.

\textbf{\ac{imu} preintegration.}
Replacing per-sample gyro/accel factors with Forster
TRO-2017~\cite{forster2017manifold} preintegrated factors at 100\,Hz
reduces constraint density from 8:1 (samples per knot) to 1:1,
degrading orientation \ac{rmse} from 1.02\degree\ to 6.40\degree\ and
position \ac{rmse} from 0.169\,m to 0.288\,m on slow racing.
The smaller Jacobian iterates faster, but the accuracy loss is
unacceptable.

\textbf{GNC (Graduated Non-Convexity).}
Geman-McClure loss annealing~\cite{yang2020gnc} on radar residuals
improved results only when extrinsics were incorrect (masking
calibration errors as outliers); with correct calibration it degraded
both bags and added $3\times$ runtime.

\textbf{Full extrinsic optimization.}
The C++ solver exposes only the 1-DoF pitch offset $\delta_\text{pitch}$
(\autoref{sec:state}): Doppler cannot observe roll or yaw, and an earlier
Python solver that freed all three angles let roll and yaw drift
5--7\degree\ into the unobservable modes as a residual trash can,
corrupting orientation by 6--8\degree.  The C++ design avoids this by
construction.
